---
title: "ADMIXTURE Analysis - All K Values"
output: html_document
date: "2024-11-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(root.dir = "/Volumes/data/popgen-new/admixture/k-sweep")
setwd("/Volumes/data/popgen-new/admixture/k-sweep")
```

```{r load-libraries}
# Load necessary libraries
library(ggplot2)
library(reshape2)
library(dplyr)
library(gridExtra)
library(grid)
library(cols4all)
```

```{r}
c4a_gui()
```


```{r read-sample-names}
# Read sample names from the file (used for all k values)
# Suppress warning about incomplete final line
sample_names <- readLines("sample_identifiers.txt", warn = FALSE)
```

## Define Colorblind-Friendly Palette

```{r define-colors}
# Using the 'friendly7' palette from cols4all
# Get the friendly7 palette colors
friendly7_palette <- c4a("friendly7", 7)

# Display the palette to verify colors
print("Friendly7 palette colors:")
print(friendly7_palette)

# Assign colors to populations based on desired characteristics:
# HEMP: light blue
# 'indica' MJ: yellow-orange-tan
# 'sativa' MJ: green
# HC_HEMP: red-orange-brown
cb_colors <- list(
  hemp = friendly7_palette[4],           # light blue for HEMP
  indica_mj = friendly7_palette[2],      # yellow-orange-tan for 'indica' MJ
  sativa_mj = friendly7_palette[3],      # green for 'sativa' MJ  
  hc_hemp_1 = friendly7_palette[1],      # red-orange-brown for HC_HEMP_1
  hc_hemp_2 = friendly7_palette[5]       # red-orange-brown variant for HC_HEMP_2
)

# Display assigned colors
print("Assigned colors:")
print(cb_colors)
```

## Process All K Values

```{r process-all-k}
# Function to process Q matrix for any k value
process_q_matrix <- function(k_value, sample_names) {
  # Read the Q matrix
  filename <- paste0("adm-k", k_value, ".Q")
  q_matrix <- read.table(filename, header = FALSE)
  
  # Round to 3 decimal places
q_matrix <- q_matrix %>% mutate(across(starts_with("V"), ~round(., 3)))
  
  # Ensure the number of sample names matches
  if (length(sample_names) != nrow(q_matrix)) {
    stop(paste("The number of sample names does not match the number of samples in K=", 
               k_value, " Q matrix."))
  }
  
  # Add sample names to the Q matrix
  q_matrix$Sample <- sample_names
  
  # Reshape to long format
  q_matrix_long <- melt(q_matrix, id.vars = "Sample", 
                       variable.name = "Population", value.name = "Proportion")
  
  # Calculate proportions for each population
  for (i in 1:k_value) {
    pop_name <- paste0("V", i)
    prop_name <- paste0("V", i, "_Percentage")
    
    prop_data <- q_matrix_long %>%
      filter(Population == pop_name) %>%
      dplyr::select(Sample, !!prop_name := Proportion)
    
    q_matrix_long <- merge(q_matrix_long, prop_data, by = "Sample")
  }
  
  return(q_matrix_long)
}

# Process all k values
q_k2 <- process_q_matrix(2, sample_names)
q_k3 <- process_q_matrix(3, sample_names)
q_k4 <- process_q_matrix(4, sample_names)
q_k5 <- process_q_matrix(5, sample_names)
```

## Define Sorting Logic and draw the plots

```{r define-sorting}
# K=2: V1=MJ, V2=HEMP
# Sort: desc(V1_MJ) - high MJ at bottom, high HEMP at top
sorted_k2 <- q_k2 %>%
  distinct(Sample, V1_Percentage, V2_Percentage) %>%
  arrange(desc(V1_Percentage), V2_Percentage)

q_k2$Sample <- factor(q_k2$Sample, levels = sorted_k2$Sample)

# K=3: V1=HEMP, V2=HC_HEMP, V3='indica' MJ
# Sort: desc(V3_indica) - high 'indica' MJ at bottom, high HEMP at top
sorted_k3 <- q_k3 %>%
  distinct(Sample, V1_Percentage, V2_Percentage, V3_Percentage) %>%
  arrange(desc(V3_Percentage), V1_Percentage, desc(V2_Percentage))

q_k3$Sample <- factor(q_k3$Sample, levels = sorted_k3$Sample)
q_k3$Population <- factor(q_k3$Population, levels = c("V3", "V2", "V1"))

# K=4: V1='indica' MJ, V2='sativa' MJ, V3=HEMP, V4=HC_HEMP
# Sort: desc(V1_indica) - high 'indica' MJ at bottom, high HEMP at top
sorted_k4 <- q_k4 %>%
  distinct(Sample, V1_Percentage, V2_Percentage, V3_Percentage, V4_Percentage) %>%
  arrange(desc(V1_Percentage), V3_Percentage, desc(V2_Percentage))

q_k4$Sample <- factor(q_k4$Sample, levels = sorted_k4$Sample)

# K=5: V1='sativa' MJ, V2=HEMP, V3='indica' MJ, V4=HC_HEMP_1, V5=HC_HEMP_2
# Sort: desc(V3_indica) - high 'indica' MJ at bottom, high HEMP at top
# Try this to get HEMP at top and MJ at bottom
sorted_k5 <- q_k5 %>%
  distinct(Sample, V1_Percentage, V2_Percentage, V3_Percentage, V4_Percentage, V5_Percentage) %>%
  arrange(desc(V3_Percentage), V2_Percentage, desc(V1_Percentage), desc(V4_Percentage), desc(V5_Percentage))

# Sort by HEMP desc, then MJ
q_k5$Sample <- factor(q_k5$Sample, levels = sorted_k5$Sample)

#Sort indica to right, and hemp to left
q_k5$Population <- factor(q_k5$Population, levels = c("V3", "V1", "V5", "V4", "V2"))

## Create Plots for All K Values
# K=2 Plot
plot_k2 <- ggplot(q_k2, aes(y = Sample, x = Proportion, fill = Population)) + 
  geom_bar(stat = "identity") + 
  scale_fill_manual(name = "K=2", 
                    values = c("V1" = cb_colors$indica_mj, 
                              "V2" = cb_colors$hemp),
                    breaks = c("V2", "V1"),
                    labels = c("HEMP", "MJ")) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 10, face = "bold"),
    axis.text.y = element_text(size = 6, hjust = 1),
    axis.text.x = element_text(size = 8),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    legend.position = "right",
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 9, face = "bold"),
    plot.margin = margin(5, 5, 5, 5)
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = "Sample ID")

print(plot_k2)

# K=3 Plot
plot_k3 <- ggplot(q_k3, aes(y = Sample, x = Proportion, fill = Population)) + 
  geom_bar(stat = "identity") + 
  scale_fill_manual(name = "K=3", 
                    values = c("V1" = cb_colors$hemp, 
                              "V2" = cb_colors$hc_hemp_1, 
                              "V3" = cb_colors$indica_mj),
                    breaks = c("V1", "V2", "V3"),
                    labels = c("HEMP", "HC_HEMP", "MJ")) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_text(size = 6, hjust = 1),
    axis.text.x = element_text(size = 8),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    legend.position = "right",
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 9, face = "bold"),
    plot.margin = margin(5, 5, 5, 5)
  ) +
  scale_x_continuous(expand = c(0, 0))

print(plot_k3)

# K=4 Plot
plot_k4 <- ggplot(q_k4, aes(y = Sample, x = Proportion, fill = Population)) + 
  geom_bar(stat = "identity") + 
  scale_fill_manual(name = "K=4", 
                    values = c("V1" = cb_colors$indica_mj, 
                              "V2" = cb_colors$sativa_mj, 
                              "V3" = cb_colors$hemp, 
                              "V4" = cb_colors$hc_hemp_1),
                    breaks = c("V3", "V4", "V2", "V1"),
                    labels = c("HEMP", "HC_HEMP", "'sativa' MJ", "'indica' MJ")) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_text(size = 6, hjust = 1),
    axis.text.x = element_text(size = 8),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    legend.position = "right",
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 9, face = "bold"),
    plot.margin = margin(5, 5, 5, 5)
  ) +
  scale_x_continuous(expand = c(0, 0))

print(plot_k4)

# K=5 Plot
plot_k5 <- ggplot(q_k5, aes(y = Sample, x = Proportion, fill = Population)) + 
  geom_bar(stat = "identity") + 
  scale_fill_manual(name = "K=5", 
                    values = c("V1" = cb_colors$sativa_mj,
                              "V2" = cb_colors$hemp, 
                              "V3" = cb_colors$indica_mj, 
                              "V4" = cb_colors$hc_hemp_2, 
                              "V5" = cb_colors$hc_hemp_1),
                    breaks = c("V2", "V5", "V4", "V1", "V3"),
                    labels = c("HEMP", "HC_HEMP_1", "HC_HEMP_2", "'sativa' MJ", "'indica' MJ")) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_text(size = 6, hjust = 1),
    axis.text.x = element_text(size = 8),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    legend.position = "right",
    legend.text = element_text(size = 8),
    legend.title = element_text(size = 9, face = "bold"),
    plot.margin = margin(5, 5, 5, 5)
  ) +
  scale_x_continuous(expand = c(0, 0))

print(plot_k5)
```

## Main Figure (K=4 only)

```{r save-main-figure}
# Save the main figure for K=4 with predetermined size
ggsave("admixture-k4-main.png", 
       plot = plot_k4, 
       width = 6.27, 
       height = 9.2, 
       units = "in", 
       dpi = 600)

# Save as PDF (A4 portrait size)
ggsave("admixture-k4-main.pdf", 
       plot = plot_k4, 
       width = 6.27, 
       height = 9.2, 
       units = "in",
       dpi = 600)

print("Main figure (K=4) saved successfully!")
```

## Supplemental Figure (K=2, K=3, K=5)

```{r supplemental-figure, fig.width=21, fig.height=29.7}
# A4 portrait size in cm: 21 x 29.7
# Combine K=2, K=3, and K=5 in three columns
supplemental_plot <- grid.arrange(
  plot_k2, 
  plot_k3, 
  plot_k5,
  ncol = 3,
  widths = c(1, 1, 1),
  bottom = textGrob("Ancestry Proportion", 
                    gp = gpar(fontsize = 10, fontface = "bold"))
)

print(supplemental_plot)

# Save as PNG (A4 portrait size at 600 DPI)
ggsave("admixture-supplemental.png", 
       plot = supplemental_plot, 
       width = 21, 
       height = 29.7, 
       units = "cm", 
       dpi = 600)

# Save as PDF (A4 portrait size)
ggsave("admixture-supplemental.pdf", 
       plot = supplemental_plot, 
       width = 21, 
       height = 29.7, 
       units = "cm",
       dpi = 600)

print("Supplemental figure saved successfully!")
```



## Color Palette Information

The plots use the 'friendly7' colorblind-friendly palette from the cols4all package with the following assignments:

- **HEMP**: Light blue (friendly7[3])
- **'indica' MJ**: Yellow-orange-tan (friendly7[4])
- **'sativa' MJ**: Green (friendly7[2])
- **HC_HEMP / HC_HEMP_1**: Red-orange-brown (friendly7[1])
- **HC_HEMP_2**: Red-orange-brown variant (friendly7[5])

These colors are distinguishable for individuals with various types of color vision deficiency.

## Sorting Logic

All plots use consistent sorting logic:
- Samples with high 'indica' MJ content appear at the **bottom**
- Samples with high HEMP content appear at the **top**
- For samples with similar primary ancestry, secondary ancestries are used for fine sorting
