R---
title: "PCA Test - SNP Data"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

# From https://towardsdatascience.com/10-tips-for-choosing-the-optimal-number-of-clusters-277e93d72d92

# Data from PLINK2 PCA - SNP Analysis

# First, let's load the required packages.


```{r}
library(tidyverse)
library(magrittr)
library(cluster)
library(cluster.datasets)
library(cowplot)
library(NbClust)
library(clValid)
library(ggfortify)
library(clustree)
library(dendextend)
library(factoextra)
library(FactoMineR)
library(corrplot)
library(GGally)
library(ggiraphExtra)
library(knitr)
library(kableExtra)
```

# Set WD
```{r}
knitr::opts_chunk$set(root.dir = "/Volumes/data/lh/popgen/nlr-pca/mpileup")
```
# Load the PCA results from PLINK2


```{r}
eigenvec_data <- read.table("/Volumes/data/lh/popgen/nlr-pca/mpileup/nlr-pca-85-plink-test-10-renamed.eigenvec.tsv", header = FALSE)
```

# Assign appropriate column names (FID, IID, PC1, PC2, etc.)
```{r}
colnames(eigenvec_data) <- c("FID", "IID", paste0("PC", 1:(ncol(eigenvec_data) - 2)))
```

Glimpse the data

```{r}
glimpse(eigenvec_data)
```




# Visualize the first two principal components

```{r}
ggplot(eigenvec_data, aes(x = PC1, y = PC2)) +
    geom_point() +
    labs(title = "PCA of SNP data", x = "PC1", y = "PC2") +
    theme_minimal()
```


# Set the number of clusters

```{r}
k <- 2 

# Perform k-means clustering on the PCA scores (using PC1 and PC2 as example)
set.seed(123)
kmeans_result <- kmeans(eigenvec_data[, 3:ncol(eigenvec_data)], centers = k, nstart = 30)

# Visualize the clustering result
ggplot(eigenvec_data, aes(x = PC1, y = PC2, color = factor(kmeans_result$cluster))) +
  geom_point() +
  labs(title = paste("k-means Clustering (k =", k, ")"),
       x = "PC1: 8.13%",
       y = "PC2: 6.45%",
       color = "Group") +
  scale_color_discrete(labels = c("MJ", "HEMP")) +  # rename legend entries
  theme_minimal() +
  theme(
    legend.text = element_text(size = 12),  # increase legend text size
    axis.title = element_text(size = 12)    # increase axis title size
  )

print(q)
```


# Save the plot as a PNG
```{r}

ggsave("nlr-pca-k2.png", plot = q, width = 12, height = 12, units = "in", dpi = 300, bg = "white")

```


# Evaluating stability of clusters using Clustree for different k values (1 to 7)

```{r}
tmp <- list()  # Initialize an empty list to store the kmeans results
for (k in 1:7) {
  tmp[[k]] <- kmeans(eigenvec_data[, 3:ncol(eigenvec_data)], centers = k, nstart = 30)$cluster
}
df <- data.frame(tmp)  # Convert the list into a data frame
```


# Add a prefix to the column names

```{r}
colnames(df) <- seq(1:7)
colnames(df) <- paste0("k", colnames(df))
```

# Get individual PCA

```{r}
df.pca <- prcomp(df, center = TRUE, scale. = FALSE)
ind.coord <- df.pca$x
ind.coord <- ind.coord[, 1:2]
df <- bind_cols(as.data.frame(df), as.data.frame(ind.coord))
```

# Visualize with Clustree

```{r}
clustree(df, prefix = "k")

```


# Hierarchical clustering using Euclidean distance

```{r}
dist_matrix <- dist(eigenvec_data[, 3:ncol(eigenvec_data)], method = "euclidean")
hc <- hclust(dist_matrix, method = "ward.D2")
```

```{r}
labels <- eigenvec_data$IID  # Replace with the appropriate column name if needed

```

# Plot the dendrogram

```{r}
plot(hc, labels = labels)
rect.hclust(hc, k = 4, border = 2:5)  # Cut tree into 4 groups
```
# make it bigger

```{r}
# Perform hierarchical clustering
dist_matrix <- dist(eigenvec_data[, 3:ncol(eigenvec_data)])  # Adjust based on your data
hc <- hclust(dist_matrix, method = "ward.D2")

# Open a new plotting window with a larger size (make it 3x bigger)
dev.new(width = 15, height = 15)  # Adjust width and height to increase plot size

# Adjust plot margins (if necessary) to accommodate larger plot
par(mar = c(8, 8, 4, 2))  # Increase margins

# Plot the dendrogram with custom labels, and make the text smaller (cex < 1)
plot(hc, labels = labels, cex = 0.5)  # 'cex = 0.5' makes the text smaller

# Cut the tree into 4 groups
rect.hclust(hc, k = 2, border = 2:5)


```

