---
title: "PCA Test - SNP Data"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

# From https://towardsdatascience.com/10-tips-for-choosing-the-optimal-number-of-clusters-277e93d72d92

# Data from PLINK2 PCA - SNP Analysis

# First, let's load the required packages.


```{r}
library(tidyverse)

library(cluster)
library(cluster.datasets)
library(cowplot)
library(NbClust)
library(clValid)
library(ggfortify)
library(clustree)
library(dendextend)
library(factoextra)
library(FactoMineR)
library(corrplot)
library(GGally)
library(knitr)
library(kableExtra)
library(ggiraphExtra)
library(magrittr)
library(cols4all)
```


## Define Colorblind-Friendly Palette
```{r define-colors}
# Using the 'friendly7' palette from cols4all
# Get the friendly7 palette colors
friendly7_palette <- c4a("friendly7", 7)

# Display the palette to verify colors
print("Friendly7 palette colors:")
print(friendly7_palette)

# Assign colors to populations based on desired characteristics:
# Group 1: HC_HEMP_2 - red-orange-brown variant
# Group 2: MJ - yellow-orange-tan for indica MJ / green for sativa MJ
# Group 3: HEMP - light blue
# Group 4: HC_HEMP_1 - red-orange-brown

cb_colors <- list(
  hc_hemp_2 = friendly7_palette[5],      # red-orange-brown variant for HC_HEMP_2 (Group 1)
  mj = friendly7_palette[3],             # green for MJ (Group 2)
  hemp = friendly7_palette[4],           # light blue for HEMP (Group 3)
  hc_hemp_1 = friendly7_palette[1]       # red-orange-brown for HC_HEMP_1 (Group 4)
)

# Display assigned colors
print("Assigned colors:")
print(cb_colors)

# Create a named vector for k=4 clustering
# Mapping cluster numbers to group labels
k4_colors <- c(
  "1" = cb_colors$hc_hemp_2,   # Group 1 = HC_HEMP_2
  "2" = cb_colors$mj,          # Group 2 = MJ
  "3" = cb_colors$hemp,        # Group 3 = HEMP
  "4" = cb_colors$hc_hemp_1    # Group 4 = HC_HEMP_1
)

k4_labels <- c(
  "1" = "HC_HEMP_2",
  "2" = "MJ",
  "3" = "HEMP",
  "4" = "HC_HEMP_1"
)
```

```{r define-colors}
# Using the 'friendly7' palette from cols4all
# Get the friendly7 palette colors
friendly7_palette <- c4a("friendly7", 7)

# Display the palette to verify colors
print("Friendly7 palette colors:")
print(friendly7_palette)

# Assign colors to populations based on desired characteristics:
# Group 1: HC_HEMP_2 - red-orange-brown variant
# Group 2: MJ - yellow-orange-tan for indica MJ / green for sativa MJ
# Group 3: HEMP - light blue (darkened)
# Group 4: HC_HEMP_1 - red-orange-brown

# Function to darken a color
darken_color <- function(color, factor = 0.9) {
  rgb_vals <- col2rgb(color) / 255
  darkened <- rgb_vals * factor
  rgb(darkened[1], darkened[2], darkened[3])
}

cb_colors <- list(
  hc_hemp_2 = friendly7_palette[5],      # red-orange-brown variant for HC_HEMP_2 (Group 1)
  mj = friendly7_palette[3],             # green for MJ (Group 2)
  hemp = darken_color(friendly7_palette[4]),  # darkened light blue for HEMP (Group 3)
  hc_hemp_1 = friendly7_palette[1]       # red-orange-brown for HC_HEMP_1 (Group 4)
)

# Display assigned colors
print("Assigned colors:")
print(cb_colors)

# Create a named vector for k=4 clustering
# Mapping cluster numbers to group labels
k4_colors <- c(
  "1" = cb_colors$hc_hemp_2,   # Group 1 = HC_HEMP_2
  "2" = cb_colors$mj,          # Group 2 = MJ
  "3" = cb_colors$hemp,        # Group 3 = HEMP
  "4" = cb_colors$hc_hemp_1    # Group 4 = HC_HEMP_1
)

k4_labels <- c(
  "1" = "HC_HEMP_2",
  "2" = "MJ",
  "3" = "HEMP",
  "4" = "HC_HEMP_1"
)
```

# I've added a `darken_color()` function that multiplies the RGB values by a factor (0.75 in this case, which makes it 25% darker). You can adjust the factor if you want it darker (use a smaller number like 0.7) or less dark (use a larger number like 0.8).

# Prepare the environment 
```{r}
#Set WD
knitr::opts_chunk$set(root.dir = "/Volumes/18tb/lh/align/popgen-85/all/pca")
setwd("/Volumes/18tb/lh/align/popgen-85/all/pca")
# Load the PCA results from PLINK2
eigenvec_data <- read.table("joint-all-plink-test-lbl.eigenvec", header = FALSE)
# Assign appropriate column names (FID, IID, PC1, PC2, etc.)
colnames(eigenvec_data) <- c("FID", "IID", paste0("PC", 1:(ncol(eigenvec_data) - 2)))
```


# Visualize the essentials
```{r}
#Glimpse the data
glimpse(eigenvec_data)
# Visualize the first two principal components
ggplot(eigenvec_data, aes(x = PC1, y = PC2)) +
    geom_point() +
    labs(title = "PCA of SNP data", x = "PC1", y = "PC2") +
    theme_minimal()
```



# K-means clustering

```{r}
#Set the number of clusters

k <- 4 

# Extract only numeric PC columns for clustering
pc_cols <- grep("^PC", names(eigenvec_data), value = TRUE)
pca_data <- eigenvec_data[, pc_cols]

# Perform k-means clustering on the PCA scores
set.seed(123)
kmeans_result <- kmeans(pca_data, centers = k, nstart = 30)

# Create a labeled cluster column
eigenvec_data$cluster_label <- factor(kmeans_result$cluster, 
                                       levels = c("1", "2", "3", "4"),
                                       labels = c("HC_HEMP_2", "MJ", "HEMP", "HC_HEMP_1"))

# Reorder factor levels for legend display: HEMP, HC_HEMP_1, HC_HEMP_2, MJ
eigenvec_data$cluster_label <- factor(eigenvec_data$cluster_label,
                                       levels = c("HEMP", "HC_HEMP_1", "HC_HEMP_2", "MJ"))

# Create the plot with smaller text and larger plot size, using labels in place of dots
p <- ggplot(eigenvec_data, aes(x = PC1, y = PC2, color = cluster_label)) +
  geom_text(aes(label = FID), vjust = 1, hjust = 1, size = 3) +  # Use labels instead of points
  labs(title = paste("k-means Clustering (k =", k, ")"), x = "PC1: 9.1%", y = "PC2: 7.6%", color = "Group") +
  scale_color_manual(values = c("HEMP" = cb_colors$hemp,
                                 "HC_HEMP_1" = cb_colors$hc_hemp_1,
                                 "HC_HEMP_2" = cb_colors$hc_hemp_2,
                                 "MJ" = cb_colors$mj)) +
  theme_minimal() +
  theme(
    text = element_text(size = 8),  # Make title and axis labels smaller
    plot.title = element_text(size = 10)  # Adjust title size specifically
  )

# Print the plot
print(p)
```



# Save the plot as a PNG file

```{r}
ggsave("k4-lbl.png", 
         plot = p, width = 21, 
         height = 21,
        units = "cm", 
       dpi = 600)
```


# K-means clustering with k = 2, 3, 5

```{r}

library(gridExtra)
library(grid)

# Extract PCA data (only numeric PC columns, exclude cluster_label if present)
# Get column names that start with "PC"
pc_cols <- grep("^PC", names(eigenvec_data), value = TRUE)
pca_data <- eigenvec_data[, pc_cols]

k_values <- c(2, 3, 5)
kmeans_plot_list <- list()
panel_labels <- c("a) k=2", "b) k=3", "c) k=5")

for (i in seq_along(k_values)) {
  k_val <- k_values[i]
  
  # Run k-means with the current k value
  set.seed(123)
  kmeans_result <- kmeans(pca_data, centers = k_val, nstart = 30)
  
  # Print summary
  cat("\nk =", k_val, "\n")
  cat("Total within-cluster sum of squares:", kmeans_result$tot.withinss, "\n")
  print(table(kmeans_result$cluster))
  
  # Store clustering results in a temporary data frame to force evaluation
  plot_data <- eigenvec_data
  plot_data$cluster <- factor(kmeans_result$cluster)
  
  # Use specific colors and labels based on k value
  if (k_val == 2) {
    # k=2: group 1 is MJ/HEMP (green), group 2 is HC_HEMP (red-orange)
    # Legend order: HC_HEMP, MJ/HEMP
    plot_data$cluster <- factor(plot_data$cluster, levels = c("2", "1"))
    current_colors <- setNames(c(cb_colors$hc_hemp_1, cb_colors$mj),
                                c("2", "1"))
    current_labels <- setNames(c("HC_HEMP", "MJ/HEMP"), c("2", "1"))
  } else if (k_val == 3) {
    # k=3: group 1 is HC_HEMP, group 2 is HEMP, group 3 is MJ
    # Legend order: HEMP, HC_HEMP, MJ
    plot_data$cluster <- factor(plot_data$cluster, levels = c("2", "1", "3"))
    current_colors <- setNames(c(cb_colors$hemp, cb_colors$hc_hemp_1, cb_colors$mj),
                                c("2", "1", "3"))
    current_labels <- setNames(c("HEMP", "HC_HEMP", "MJ"), c("2", "1", "3"))
  } else if (k_val == 5) {
    # k=5: group 1 is 'indica' MJ, group 2 is HC_HEMP_2, group 3 is 'sativa' MJ, group 4 is HC_HEMP_1, group 5 is HEMP
    # Legend order: HEMP, HC_HEMP_1, HC_HEMP_2, 'sativa' MJ, 'indica' MJ
    plot_data$cluster <- factor(plot_data$cluster, levels = c("5", "4", "2", "3", "1"))
    current_colors <- setNames(c(cb_colors$hemp, cb_colors$hc_hemp_1, cb_colors$hc_hemp_2, friendly7_palette[2], friendly7_palette[3]),
                                c("5", "4", "2", "3", "1"))
    current_labels <- setNames(c("HEMP", "HC_HEMP_1", "HC_HEMP_2", "'sativa' MJ", "'indica' MJ"), c("5", "4", "2", "3", "1"))
  } else {
    current_colors <- setNames(friendly7_palette[1:k_val], levels(plot_data$cluster))
    current_labels <- NULL
  }
  
  p <- ggplot(plot_data, aes(x = PC1, y = PC2, color = cluster)) +
    geom_text(aes(label = FID), vjust = 1, hjust = 1, size = 2.5) +
    labs(title = panel_labels[i],
         x = "PC1: 9.1%", 
         y = "PC2: 7.6%", 
         color = "Group") +
    scale_color_manual(values = current_colors, labels = current_labels) +
    theme_minimal() +
    theme(
      text = element_text(size = 10),
      plot.title = element_text(size = 11, face = "bold"),
      legend.position = "right"
    )
  
  # Force evaluation by printing
  print(p)
  
  # Store the plot
  kmeans_plot_list[[i]] <- p
}

# Arrange all 4 plots in a 2x2 grid and save as A4 PNG
kmeans_combined <- grid.arrange(grobs = kmeans_plot_list, ncol = 2)

# Save as A4-sized PNG (21 x 29.7 cm)
ggsave("kmeans-clustering-k2-k5.png", 
       plot = kmeans_combined, 
       width = 21, 
       height = 29.7, 
       units = "cm", 
       dpi = 600, 
       bg = "white")

cat("\nK-means clustering plot saved as 'kmeans-clustering-k2-k5.png'\n")
```

# Partitioning around medioids (PAM), k=2 to 5


```{r}
# PAM clustering with k from 2 to 5
library(cluster)
library(gridExtra)
library(grid)

# Extract PCA data (only numeric PC columns, exclude cluster_label if present)
# Get column names that start with "PC"
pc_cols <- grep("^PC", names(eigenvec_data), value = TRUE)
pca_data <- eigenvec_data[, pc_cols]

k_values <- 2:5
pam_plot_list <- list()
panel_labels <- c("a) k=2", "b) k=3", "c) k=4", "d) k=5")

for (i in seq_along(k_values)) {
  k_val <- k_values[i]
  
  # Run PAM with the current k value
  pam_result <- pam(pca_data, k = k_val)
  
  # Print silhouette score
  cat("\nk =", k_val, "- Average Silhouette Width:", pam_result$silinfo$avg.width, "\n")
  cat("Number of clusters:", length(unique(pam_result$clustering)), "\n")
  print(table(pam_result$clustering))
  
  # Store clustering results in a temporary data frame to force evaluation
  plot_data <- eigenvec_data
  plot_data$cluster <- factor(pam_result$clustering)
  
  # Use specific colors and labels based on k value
  if (k_val == 2) {
    # k=2: group 1 is MJ, group 2 is HEMP
    current_colors <- setNames(c(cb_colors$mj, cb_colors$hemp),
                                c("1", "2"))
    current_labels <- setNames(c("MJ", "HEMP"), c("1", "2"))
  } else if (k_val == 3) {
    # k=3: group 1 is MJ, group 2 is HEMP, group 3 is HC_HEMP
    current_colors <- setNames(c(cb_colors$mj, cb_colors$hemp, cb_colors$hc_hemp_1),
                                c("1", "2", "3"))
    current_labels <- setNames(c("MJ", "HEMP", "HC_HEMP"), c("1", "2", "3"))
  } else if (k_val == 4) {
    # k=4: group 1 is MJ, group 2 is HEMP, group 3 is HC_HEMP_1, group 4 is HC_HEMP_2
    current_colors <- setNames(c(cb_colors$mj, cb_colors$hemp, cb_colors$hc_hemp_1, cb_colors$hc_hemp_2),
                                c("1", "2", "3", "4"))
    current_labels <- setNames(c("MJ", "HEMP", "HC_HEMP_1", "HC_HEMP_2"), c("1", "2", "3", "4"))
  } else if (k_val == 5) {
    # k=5: group 1 is MJ, group 2 is HEMP, group 3 is HC_HEMP_1, group 4 is HC_HEMP_2, group 5 is HC_HEMP_3
    current_colors <- setNames(c(cb_colors$mj, cb_colors$hemp, cb_colors$hc_hemp_1, cb_colors$hc_hemp_2, friendly7_palette[6]),
                                c("1", "2", "3", "4", "5"))
    current_labels <- setNames(c("MJ", "HEMP", "HC_HEMP_1", "HC_HEMP_2", "HC_HEMP_3"), c("1", "2", "3", "4", "5"))
  } else {
    current_colors <- setNames(friendly7_palette[1:k_val], levels(plot_data$cluster))
    current_labels <- NULL
  }
  
  p <- ggplot(plot_data, aes(x = PC1, y = PC2, color = cluster)) +
    geom_text(aes(label = FID), vjust = 1, hjust = 1, size = 2.5) +
    labs(title = panel_labels[i], 
         x = "PC1: 9.1%", 
         y = "PC2: 7.6%", 
         color = "Group") +
    scale_color_manual(values = current_colors, labels = current_labels) +
    theme_minimal() +
    theme(
      text = element_text(size = 10),
      plot.title = element_text(size = 11, face = "bold"),
      legend.position = "right"
    )
  
  # Force evaluation by printing (this ensures the plot is created now)
  print(p)
  
  # Store the plot
  pam_plot_list[[i]] <- p
}

# Arrange all 4 plots in a 2x2 grid and save as A4 PNG
pam_combined <- grid.arrange(grobs = pam_plot_list, ncol = 2)

# Save as A4-sized PNG (21 x 29.7 cm)
ggsave("pam-clustering-k2-k5.png", 
       plot = pam_combined, 
       width = 21, 
       height = 29.7, 
       units = "cm", 
       dpi = 600, 
       bg = "white")

cat("\nPAM clustering plot saved as 'pam-clustering-k2-k5.png'\n")
```
```{r}
# PAM clustering with k from 2 to 5
library(cluster)
library(gridExtra)
library(grid)

# Extract PCA data (only numeric PC columns, exclude cluster_label if present)
# Get column names that start with "PC"
pc_cols <- grep("^PC", names(eigenvec_data), value = TRUE)
pca_data <- eigenvec_data[, pc_cols]

k_values <- 2:5
pam_plot_list <- list()
panel_labels <- c("a) k=2", "b) k=3", "c) k=4", "d) k=5")

for (i in seq_along(k_values)) {
  k_val <- k_values[i]
  
  # Run PAM with the current k value
  pam_result <- pam(pca_data, k = k_val)
  
  # Print silhouette score
  cat("\nk =", k_val, "- Average Silhouette Width:", pam_result$silinfo$avg.width, "\n")
  cat("Number of clusters:", length(unique(pam_result$clustering)), "\n")
  print(table(pam_result$clustering))
  
  # Store clustering results in a temporary data frame to force evaluation
  plot_data <- eigenvec_data
  plot_data$cluster <- factor(pam_result$clustering)
  
  # Use specific colors and labels based on k value
  if (k_val == 2) {
    # k=2: group 1 is MJ, group 2 is HEMP
    # Legend order: HEMP, MJ
    plot_data$cluster <- factor(plot_data$cluster, levels = c("2", "1"))
    current_colors <- setNames(c(cb_colors$hemp, cb_colors$mj),
                                c("2", "1"))
    current_labels <- setNames(c("HEMP", "MJ"), c("2", "1"))
  } else if (k_val == 3) {
    # k=3: group 1 is MJ, group 2 is HEMP, group 3 is HC_HEMP
    # Legend order: HEMP, HC_HEMP, MJ
    plot_data$cluster <- factor(plot_data$cluster, levels = c("2", "3", "1"))
    current_colors <- setNames(c(cb_colors$hemp, cb_colors$hc_hemp_1, cb_colors$mj),
                                c("2", "3", "1"))
    current_labels <- setNames(c("HEMP", "HC_HEMP", "MJ"), c("2", "3", "1"))
  } else if (k_val == 4) {
    # k=4: group 1 is MJ, group 2 is HEMP, group 3 is HC_HEMP_1, group 4 is HC_HEMP_2
    # Legend order: HEMP, HC_HEMP_1, HC_HEMP_2, MJ
    plot_data$cluster <- factor(plot_data$cluster, levels = c("2", "3", "4", "1"))
    current_colors <- setNames(c(cb_colors$hemp, cb_colors$hc_hemp_1, cb_colors$hc_hemp_2, cb_colors$mj),
                                c("2", "3", "4", "1"))
    current_labels <- setNames(c("HEMP", "HC_HEMP_1", "HC_HEMP_2", "MJ"), c("2", "3", "4", "1"))
  } else if (k_val == 5) {
    # k=5: group 1 is MJ, group 2 is HEMP, group 3 is HC_HEMP_1, group 4 is HC_HEMP_2, group 5 is HC_HEMP_3
    # Legend order: HEMP, HC_HEMP_1, HC_HEMP_2, HC_HEMP_3, MJ
    plot_data$cluster <- factor(plot_data$cluster, levels = c("2", "3", "4", "5", "1"))
    current_colors <- setNames(c(cb_colors$hemp, cb_colors$hc_hemp_1, cb_colors$hc_hemp_2, friendly7_palette[6], cb_colors$mj),
                                c("2", "3", "4", "5", "1"))
    current_labels <- setNames(c("HEMP", "HC_HEMP_1", "HC_HEMP_2", "HC_HEMP_3", "MJ"), c("2", "3", "4", "5", "1"))
  } else {
    current_colors <- setNames(friendly7_palette[1:k_val], levels(plot_data$cluster))
    current_labels <- NULL
  }
  
  p <- ggplot(plot_data, aes(x = PC1, y = PC2, color = cluster)) +
    geom_text(aes(label = FID), vjust = 1, hjust = 1, size = 2.5) +
    labs(title = panel_labels[i], 
         x = "PC1: 9.1%", 
         y = "PC2: 7.6%", 
         color = "Group") +
    scale_color_manual(values = current_colors, labels = current_labels) +
    theme_minimal() +
    theme(
      text = element_text(size = 10),
      plot.title = element_text(size = 11, face = "bold"),
      legend.position = "right"
    )
  
  # Force evaluation by printing (this ensures the plot is created now)
  print(p)
  
  # Store the plot
  pam_plot_list[[i]] <- p
}

# Arrange all 4 plots in a 2x2 grid and save as A4 PNG
pam_combined <- grid.arrange(grobs = pam_plot_list, ncol = 2)

# Save as A4-sized PNG (21 x 29.7 cm)
ggsave("pam-clustering-k2-k5.png", 
       plot = pam_combined, 
       width = 21, 
       height = 29.7, 
       units = "cm", 
       dpi = 600, 
       bg = "white")

cat("\nPAM clustering plot saved as 'pam-clustering-k2-k5.png'\n")
```

# Gaussian Mixture Model (GMM), k-sweep from 2 to 5

```{r}
# GMM clustering with k from 2 to 5
library(mclust)
library(gridExtra)
library(grid)

# Extract PCA data (only numeric PC columns, exclude cluster_label if present)
# Get column names that start with "PC"
pc_cols <- grep("^PC", names(eigenvec_data), value = TRUE)
pca_data <- eigenvec_data[, pc_cols]

k_values <- 2:5
gmm_plot_list <- list()
panel_labels <- c("a) k=2", "b) k=3", "c) k=4", "d) k=5")

for (i in seq_along(k_values)) {
  k_val <- k_values[i]
  
  # Run GMM with the current k value
  gmm_result <- Mclust(pca_data, G = k_val)
  
  # Print summary
  cat("\nk =", k_val, "- Model:", gmm_result$modelName, "\n")
  cat("BIC:", gmm_result$bic, "\n")
  cat("Number of clusters:", gmm_result$G, "\n")
  print(table(gmm_result$classification))
  
  # Store clustering results in a temporary data frame to force evaluation
  plot_data <- eigenvec_data
  plot_data$cluster <- factor(gmm_result$classification)
  
  # Use specific colors and labels based on k value
  if (k_val == 2) {
    # k=2: group 1 is HEMP, group 2 is MJ
    # Legend order: HEMP, MJ
    plot_data$cluster <- factor(plot_data$cluster, levels = c("1", "2"))
    current_colors <- setNames(c(cb_colors$hemp, cb_colors$mj),
                                c("1", "2"))
    current_labels <- setNames(c("HEMP", "MJ"), c("1", "2"))
  } else if (k_val == 3) {
    # k=3: group 1 is HC_HEMP, group 2 is MJ, group 3 is HEMP
    # Legend order: HEMP, HC_HEMP, MJ
    plot_data$cluster <- factor(plot_data$cluster, levels = c("3", "1", "2"))
    current_colors <- setNames(c(cb_colors$hemp, cb_colors$hc_hemp_1, cb_colors$mj),
                                c("3", "1", "2"))
    current_labels <- setNames(c("HEMP", "HC_HEMP", "MJ"), c("3", "1", "2"))
  } else if (k_val == 4) {
    # k=4: group 1 is HC_HEMP_2, group 2 is MJ, group 3 is HEMP, group 4 is HC_HEMP_1
    # Legend order: HEMP, HC_HEMP_1, HC_HEMP_2, MJ
    plot_data$cluster <- factor(plot_data$cluster, levels = c("3", "4", "1", "2"))
    current_colors <- setNames(c(cb_colors$hemp, cb_colors$hc_hemp_1, cb_colors$hc_hemp_2, cb_colors$mj),
                                c("3", "4", "1", "2"))
    current_labels <- setNames(c("HEMP", "HC_HEMP_1", "HC_HEMP_2", "MJ"), c("3", "4", "1", "2"))
  } else if (k_val == 5) {
    # k=5: group 1 is 'sativa' MJ, group 2 is 'indica' MJ, group 3 is HEMP, group 4 is HC_HEMP_1, group 5 is HC_HEMP_2
    # Legend order: HEMP, HC_HEMP_1, HC_HEMP_2, 'sativa' MJ, 'indica' MJ
    plot_data$cluster <- factor(plot_data$cluster, levels = c("3", "4", "5", "1", "2"))
    current_colors <- setNames(c(cb_colors$hemp, cb_colors$hc_hemp_1, cb_colors$hc_hemp_2, friendly7_palette[2], friendly7_palette[3]),
                                c("3", "4", "5", "1", "2"))
    current_labels <- setNames(c("HEMP", "HC_HEMP_1", "HC_HEMP_2", "'sativa' MJ", "'indica' MJ"), c("3", "4", "5", "1", "2"))
  } else {
    current_colors <- setNames(friendly7_palette[1:k_val], levels(plot_data$cluster))
    current_labels <- NULL
  }
  
  p <- ggplot(plot_data, aes(x = PC1, y = PC2, color = cluster)) +
    geom_text(aes(label = FID), vjust = 1, hjust = 1, size = 2.5) +
    labs(title = panel_labels[i], 
         x = "PC1: 9.1%", 
         y = "PC2: 7.6%", 
         color = "Group") +
    scale_color_manual(values = current_colors, labels = current_labels) +
    theme_minimal() +
    theme(
      text = element_text(size = 10),
      plot.title = element_text(size = 11, face = "bold"),
      legend.position = "right"
    )
  
  # Force evaluation by printing
  print(p)
  
  # Store the plot
  gmm_plot_list[[i]] <- p
}

# Arrange all 4 plots in a 2x2 grid and save as A4 PNG
gmm_combined <- grid.arrange(grobs = gmm_plot_list, ncol = 2)

# Save as A4-sized PNG (21 x 29.7 cm)
ggsave("gmm-clustering-k2-k5.png", 
       plot = gmm_combined, 
       width = 21, 
       height = 29.7, 
       units = "cm", 
       dpi = 600, 
       bg = "white")

cat("\nGMM clustering plot saved as 'gmm-clustering-k2-k5.png'\n")
```
